% Created 2015-04-16 Thu 09:17
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage[utf8]{inputenc}
\usepackage[usenames,dvipsnames]{color}
\usepackage[backend=bibtex, style=numeric]{biblatex}
\usepackage{commath}
\usepackage{tikz}
\usetikzlibrary{shapes,backgrounds}
\usepackage{marginnote}
\usepackage{listings}
\usepackage{color}
\usepackage{enumerate}
\hypersetup{urlcolor=blue}
\hypersetup{colorlinks,urlcolor=blue}
\addbibresource{bibliography.bib}
\setlength{\parskip}{16pt plus 2pt minus 2pt}
\definecolor{codebg}{rgb}{0.96,0.99,0.8}
\definecolor{codestr}{rgb}{0.46,0.09,0.2}
\author{Oleg Sivokon}
\date{\textit{<2015-04-03 Fri>}}
\title{Assignment 12, Infinitesimal Calculus}
\hypersetup{
  pdfkeywords={Infinitesimal Calculus, Assignment, Definition of Limits},
  pdfsubject={Second asssignment in the course Infinitesimal Calculus},
  pdfcreator={Emacs 25.0.50.1 (Org mode 8.2.2)}}
\begin{document}

\maketitle
\tableofcontents


\lstset{ %
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\scriptsize,
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=false,
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  framexleftmargin=10pt,
  xleftmargin=10pt,
  framerule=0pt,
  frame=tb,                        % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stringstyle=\color{codestr},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
}

\clearpage

\section{Problems}
\label{sec-1}

\subsection{Problem 1}
\label{sec-1-1}
\begin{enumerate}
\item Prove from definition $\epsilon-N$:
\begin{equation*}
  \lim_{n \to \infty} \frac{n^2}{n^2-1} = L.
\end{equation*}

\item Prove from definition $M-N$:
\begin{equation*}
   \lim_{n \to \infty} \frac{n^2-n}{n+2} = \infty.
\end{equation*}

\item Formulate $\lim_{n \to \infty} a_n = \infty$ in terms of $M-N$ definition.
\item Prove using the formulation from (3) that:
\begin{equation*}
   \lim_{n \to \infty} (n\sqrt{2} + (-1)^n \lfloor \sqrt{n} \rfloor) = \infty.
\end{equation*}
\end{enumerate}

\subsubsection{Answer 1}
\label{sec-1-1-1}
Recall the definition:

\begin{quote}
Let $(a_n)^{\infty}_{n=1}$ be a sequence, $L$ be a real number.  The
\textbf{sequence $(a_n)$ converges to $L$} if for all $\epsilon > 0$ almost all
elements of the sequence are in the $\epsilon$-neighborhood of $L$.  In such
a case we will say $L$ is the \textbf{limit of sequence $(a_n)$}.
\end{quote}

To be honest, I don't know what this question is asking me to do.  How can
I prove something which is already given in the description of a question?
Anyways, I followed the examples in the book, and here is what I think the
answer might look like.  Sorry, if this is not what you expect.
\begin{equation*}
  \begin{aligned}
    \epsilon &> 1 \implies & \textit{From definition of reals} \\
    \epsilon &> \frac{n^2-1}{n^2-1} \implies & \textit{Provided $n \neq 1$} \\
    \epsilon' &> \frac{1}{n^2-1} \implies & \textit{From Archimedian property} \\
    \epsilon + \epsilon' &> \frac{n^2}{n^2-1} \implies
    & \textit{Provided $\epsilon'' = \epsilon' + \epsilon$} \\
    \epsilon'' &> \abs{\frac{n^2}{n^2-1} - L} & \textit{Provided $0 \leq L \leq 1$}
  \end{aligned}
\end{equation*}

Which completes the ``proof''.
\subsubsection{Answer 2}
\label{sec-1-1-2}
First, recall the definition:

\begin{quote}
We will say that the sequence $(a_n)^{\infty}_{n=1}$ \textbf{tends to infinity}, and
write $\lim_{n \to \infty} a_n = \infty$, if for every real number $M$ there
exists a natural number $N$ such that for all $n > N$ it holds that $a_n > M$.
\end{quote}

We can start by finding a more convenient formulat to work with, for instance:
\begin{equation*}
  \begin{aligned}
    M &= \frac{n^2 - n}{n + 2} \\
    &= \frac{n(n - 1)}{n + 2} \\
    &= \frac{n(n + 2)}{n + 2} - \frac{3n}{n + 2} \\
    &< n - \frac{3n}{n} \\
    &= n - 3.
  \end{aligned}
\end{equation*}

Now, put $N = M + 3$. Thus for every $M$, $a_n$ is greater than $M$ whenver
$n \geq M + 3$.
\subsubsection{Answer 3}
\label{sec-1-1-3}
I'm confused by this question in the same measure I'm confused by the previous
one.  I don't know what do you want me to prove here.  There can be different
reasons for why some sequence doesn't tend to infinity.  It could be because
it is a divergent series, or because it converges to some real limit, or
because it tends to negative infinity.  I can formally negate the statement
but I don't think this negation is useful.
\begin{equation*}
  \begin{aligned}
    \lnot \forall M \in \mathbb{R} : \exists N \in \mathbb{N} :
    \forall n \in \mathbb{N} : (n > N \implies a_n > M) \\
    \textit{Alternatively:} \\
    \exists M \in \mathbb{R} : \forall N \in \mathbb{N} :
    \exists n \in \mathbb{N} : (n > N \implies a_n \leq M)
  \end{aligned}
\end{equation*}
\subsubsection{Answer 4}
\label{sec-1-1-4}
It is easy to see that the given sequence is divergent.  It has two limit
points, where one is at infinity and another one is at zero.  Since I need
to reuse the $M-N$ rule, I will restate this clame in terms of this rule.
There exists a real number $M$, such that for every natural number $N$,
there exists a natural number $n$, such that whenever $n > N$, $a_n \leq M$.
Put $M = 2$, then $a_n = 2(\sqrt{2} + \lfloor \sqrt{2} \rfloor)$.  Now, no matter
the $N$, we can always choose $n$ to be odd, which will give us:
\begin{equation*}
  \begin{aligned}
    2\sqrt{2} + \lfloor 2\sqrt{2} \rfloor \geq n\sqrt{2} - \lfloor n\sqrt{2} \rfloor \\
    \textit{For every $n > 2$} \\
    2 * 2 + \lfloor 2 * 2 \rfloor \geq n 2 - \lfloor n * 2 \rfloor \\
    \textit{By axiom of order} \\
    8 \geq 0.
  \end{aligned}
\end{equation*}
\subsection{Problem 2}
\label{sec-1-2}
Calculate the limits of the expression given above, or prove that the limtes don't
exists:

\begin{eqnarray}
  \lim_{n \to \infty} \sqrt{1 + \frac{a}{n}} \\
  \lim_{n \to \infty} \sqrt{n + a} * \sqrt{n + b} - n \\
  \lim_{n \to \infty} \frac{n^7 - 2n^4 - 1}{n^4 - 3n^6 + 7} \\
  \lim_{n \to \infty} \Big(\frac{n}{n + 1} \sum_{k=1}^n \frac{k}{k+1} \Big)
\end{eqnarray}

\subsubsection{Answer 5}
\label{sec-1-2-1}
Using ``sandwich'' rule we can show that:
\begin{equation*}
  \sqrt{1} < \sqrt{1 + \frac{a}{n}} < 1 + \frac{a}{n}
\end{equation*}

Provided $a$ is non-negative and isn't equal to $n$.  The limit of $\sqrt{1}$
is trivially 1.  The limit of $1 + \frac{a}{n}$ is the sum of the limits of
1 and $\frac{a}{n}$, where the former is the limit of a constant (1), and the
later is zero.  Thus the value is ``sandwiched'' between 1 and 1, hence it
must be 1.
\subsubsection{Answer 6}
\label{sec-1-2-2}
Let's generalize the expression to reuse the previous case:
\begin{equation*}
  \begin{aligned}
    \lim_{n \to \infty} \sqrt{n + a} * \sqrt{n + b} - n &= \\
    \lim_{n \to \infty} \sqrt{n(1 + \frac{a}{n})} * \sqrt{n(1 + \frac{b}{n})} - n &= \\
    \lim_{n \to \infty} n\sqrt{1 + \frac{a}{n}} * \sqrt{1 + \frac{b}{n}} - n &= \\
    \lim_{n \to \infty} n\Big(\sqrt{1 + \frac{a}{n}} * \sqrt{1 + \frac{b}{n}} - 1\Big)
  \end{aligned}
\end{equation*}

Now, recall that $\lim_{n \to \infty} \sqrt{1 + \frac{a}{n}} = 1$, this gives us:
\begin{equation*}
  \begin{aligned}
    \lim_{n \to \infty} n\Big(\sqrt{1 + \frac{a}{n}} * \sqrt{1 + \frac{b}{n}} - 1\Big) &= \\
    \lim_{n \to \infty} n * \lim_{n \to \infty}\sqrt{1 + \frac{a}{n}} *
    \lim_{n \to \infty}\sqrt{1 + \frac{b}{n}} - \lim_{n \to \infty} 1 &= \\
    lim_{n \to \infty} n * (1 * 1 - 1) &= \\
    lim_{n \to \infty} n * 0 &= 0.
  \end{aligned}
\end{equation*}
\subsubsection{Answer 7}
\label{sec-1-2-3}
I know that this expression tends to negative infinity as $n$ tends to infinity
simply because the numerator has a power of $n$ greater than the denominator and
the greatest term in the denominator is negative.  But I couldn't find a way
to prove this formally.
\subsubsection{Answer 8}
\label{sec-1-2-4}
I will show that the given sequence tentds to infinity.  Some notation first.
I will use $H_n$ to denote the \emph{nth} harmonic numbers.  Let's
at first, simplify the sum:
\begin{equation*}
  \begin{aligned}
    \sum_{k=1}^n \frac{k}{k + 1} &= \\
    \sum_{k=1}^n \frac{k + 1 - 1}{k + 1} &= \\
    \sum_{k=1}^n \frac{k + 1}{k + 1} - \frac{1}{k + 1} &= \\
    \sum_{k=1}^n \frac{k + 1}{k + 1} - \sum_{k=1}^n \frac{1}{k + 1} &= \\
    n - \sum_{k=1}^n \frac{1}{k + 1} &= \\
    n - \sum_{k=2}^{n+1} \frac{1}{k} &= \\
    n - H_{n+1} + 1
  \end{aligned}
\end{equation*}

Now, plug this back into our formula:
\begin{equation*}
  \begin{aligned}
    \lim_{n \to \infty} \frac{n}{n + 1} (n - H_{n+1} + 1) &= \\
    \lim_{n \to \infty} \frac{n}{n + 1} (n - H_{n+1}) + \lim_{n \to \infty}1 &= \\
    \lim_{n \to \infty} \frac{n}{n + 1} (n - H_{n+1}) + 1 &= \\
    \lim_{n \to \infty} \Big( \frac{n^2}{n + 1} - \frac{n}{n + 1} H_{n+1} \Big) + 1 &= \\
    \lim_{n \to \infty} \Big( \frac{n^2 + 1}{n + 1} - \frac{1}{n + 1} -
    \frac{n}{n + 1} H_{n+1} \Big) + 1 &= \\
    \lim_{n \to \infty} \Big( \frac{n(n + 1)}{n + 1} - \frac{1}{n + 1} -
    \frac{n}{n + 1} H_{n+1} \Big) + 1 &= \\
    \lim_{n \to \infty} \Big( n - \frac{1}{n + 1} - \frac{n}{n + 1} H_{n+1} \Big) + 1 &= \\
    \lim_{n \to \infty} n - \lim_{n \to \infty} \frac{1}{n + 1} -
    \lim_{n \to \infty} \frac{n}{n + 1} H_{n+1} + 1 &= \\
    \lim_{n \to \infty} n - 0 - \lim_{n \to \infty} 1 * H_{n+1} + 1 &= \\
    \lim_{n \to \infty} n &= \infty.
  \end{aligned}
\end{equation*}

The last step is allowed becasue for all harmonic numbers greater than one
it holds that $n > H_{n+1}$.
\subsection{Problem 3}
\label{sec-1-3}
Let $(a_n)$ be a sequence such that $\lim_{n \to \infty}(a_{n+1} - a_n) = 0$.
\begin{enumerate}
\item Prove that if there exists a constant $c > 0$ s.t. for all $n$ 
$\abs{a_n}\geq c$, then almost all elements of the sequence are either
negative or positive.
\item Show that the same doesn't hold for $\abs{a_n} > 0$.
\item Show that if for all $n$ it holds that $\abs{a_n} \geq n$, then $(a_n)$
converges in a general sense.
\end{enumerate}

\subsubsection{Answer 9}
\label{sec-1-3-1}
First, observe that the definition gives us a Cauchy sequence.  Recall that
Cauchy sequence is a sequence where intervals between subseqent members tend
to zero as $n$ tends to infinity.  That said, from the definition of the
limit, we can find $\epsilon < c$, such that almost all members of the
sequence would be in the vicinity of $c$ no farther away than
$\epsilon$. Thus, the sequence would converge to either a negative number or
a positive number, but not zero.  Since almost all members of a sequence can
be showed to be in a vicinity of a number other than zero, we have proved
the claim.
\subsubsection{Answer 10}
\label{sec-1-3-2}
The example $(a_n)$ such that doesn't match the requirement from the previous
question is, for example, $a_n = \frac{(-1)^n}{2n}$.  I.e. an alternating
series where even members are positive and odd members are negative.  The
general form of $a_{n+1} - a_n$ will thus look like this:
\begin{equation*}
  \lim_{n \to \infty}\Big(\frac{(-1)^{n+1}-2(-1)^n}{2n}\Big) = 0.
\end{equation*}


It's easy to see that this limit is zero: the nominator is a constant term
because one raised to any power is either one or negative one, while
denominator is growing with $n$.  It is also easy to see that this is
an alternating series, i.e. the sign alters with parity of $n$.
\subsubsection{Answer 11}
\label{sec-1-3-3}
Recall \ref{sec-1-3-1}, where we used the fact that $(a_n)$ is described as a
Cauchy sequence.  This is enough to show that this sequence converges in
a general sense, whether the equation $\abs{a_n} \geq n$ holds or not
is inconsequential.  Perhaps it's worth saying that the above requirement
makes it explicit that the sequence converges to infinity if any of its
members is positive, and negative infinity otherwise.
\subsection{Problem 4}
\label{sec-1-4}
Given sequences $(a_n)$ and $(b_n)$ s.t. $\lim_{n \to \infty}(a_nb_n) = 0$,
prove or disprove the following:

\begin{enumerate}
\item Either $\lim_{n \to \infty}a_n = 0$ or $\lim_{n \to \infty}b_n = 0$.
\item If $\lim_{n \to \infty}b_n = 1$, then there exists $N > 0$ s.t.
      $a_n < \frac{1}{2}$ for all $n > N$.
\item If $\lim_{n \to \infty}b_n = 1$, then $\lim_{n \to \infty}a_n = 0$.
\item If almost all members of $(b_n)$ are positive, then 
      $\lim_{n \to \infty}a_n \neq \infty$.
\item If almost all members of $(b_n)$ are positive, then 
      $\lim_{n \to \infty}a_n = 0$.
\item If there exists a constant $c > 0$ s.t. almost all $b_n \geq c$, then
      $\lim_{n \to \infty}a_n = 0$.
\end{enumerate}

\subsubsection{Answer 12}
\label{sec-1-4-1}
The limit of a product is a product of limits (their existence is
provisional).  This gives us an equation in two unknowns: $ab = 0$.
From this equation follows that either $a$ or $b$ must be zero
Thus the claim is true.
\subsubsection{Answer 13}
\label{sec-1-4-2}
As shown in \ref{sec-1-4-1}, either one (or both) limits must be zero,
thus if $\lim_{n \to \infty}a_n$ must be zero.  From definition of
the limit, there must be an $\epsilon$, such that for any $N$, almost
all $\abs{a_n - L} < \epsilon$, for $n > N$.  We can choose
$\epsilon = \frac{1}{2}$, since $L$ (the limit of $(a_n)$) is zero
the inequality becomes $\abs{a_n} < \frac{1}{2}$ since
$a_n \leq \abs{a_n} < \frac{1}{2}$ the claim is true.
\subsubsection{Answer 14}
\label{sec-1-4-3}
The claim is true. This is simply a corollary of \ref{sec-1-4-1}.
\subsubsection{Answer 16}
\label{sec-1-4-4}
This is not true.  Put $b_n = \frac{1}{n}$, the limit of $(b_n)$ is 0.  Even
if $(a_n)$ is a divergent series, for example, $a_n = \sqrt{n}$ the limit of
a product is still zero.
\subsubsection{Answer 17}
\label{sec-1-4-5}
Same example as \ref{sec-1-4-4}.  Limit of $(a_n)$ isn't zero and yet the
requirement is satisfied, thus the claim is false.
\subsubsection{Answer 18}
\label{sec-1-4-6}
Existence of lower bound greater than zero precludes $(b_n)$ from having
zero as its limit.  Assuming that there are members of a sequence which
are smaller than the lower bound is in direct contradiction with the
bound definition.  Thus the limit of $(a_n)$ must be zero (follows from
what we proved in \ref{sec-1-4-1}).  Thus the claim is true.
% Emacs 25.0.50.1 (Org mode 8.2.2)
\end{document}